---
title: "Hake eDNA Supplement 1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

load("../Stan Model Fits/_Summarized_Output/Qpcr_summaries_lat.long.smooth_Base_Var_4_10_fix_nu_T.RData")
```

## This is a list of supplemental figures needed for the eDNA paper

### Marginal smooth plots.

## Environmental DNA analysis

This section provides detailed information about the analysis of eDNA. The statistical model for eDNA is presented in the main text.

### Standards

For each qPCR plate, we included known samples with a known number of hake DNA copies to allow us to translate the observed PCR cycle at which amplification occurred to the number of copies in a sample. We present both the relationship for the PCR cycle at which amplification was observed  (Fig. \ref{fig:stand}) and the probability of observing amplification at different DNA copy numbers (Fig. \ref{fig:stand.pres}). The analysis of standard show that our qPCR analyses were sensitive to very low copy numbers and consistent across plates.
 
```{r fig.stand, echo=FALSE,fig.cap="\\label{fig:stand} qPCR standards for 33 plates with DNA copy numbers ranging from 1 to 100,000." ,out.width='100%'}
print(Output.summary.qpcr$stand.plot)
```

```{r fig.stand.pres, echo=FALSE,fig.cap="\\label{fig:stand.pres} Amplification success for qPCR standards for 33 plates with DNA copy numbers ranging from 1 to 100,000.",out.width='75%'}
print(Output.summary.qpcr$stand.plot.pres)
```

### Investigations of contamination.

In addition to field samples, we collected two kinds of control samples ship board to look for potential contamination of samples during sample processing. We found low levels of contamination in while control samples.  While 18 of 49 had hake DNA below the detection limit of 20 $copies L^{-1}$, 31 had higher concentrations. Of these 31 samples, most (21 samples) had very low concentrations $<100 copies L^{-1}$ but there was a single sample with $>$ $500$ $copies L^{-1}$ (Fig. \ref{fig:contam.hist}). From these observations, we estimated the parameters from a log-normal distribution to describe the distribution of hake DNA from contamination (Fig. \ref{fig:contam.hist}; posterior mean estimates:
$\mu$ = `r round(Output.summary.qpcr$mu_contam_out$Mean[Output.summary.qpcr$mu_contam_out$type=="liter"],3)`, $\sigma =$ `r round(Output.summary.qpcr$sigma_contam_out$Mean,3)`).This corresponds to an expected contribution of `r round(exp(Output.summary.qpcr$mu_contam_out$Mean[Output.summary.qpcr$mu_contam_out$type=="liter"] + 0.5*Output.summary.qpcr$sigma_contam_out$Mean^2),1)` copies per liter per sample. For this distribution 92% of samples would be expected to have less than 100 $copies$ $L^{-1}$ but there will be a small number of more significantly contaminated samples. We remain unsure of the ultimate source of contamination in these samples and are conducting 

To understand the potential effect of contamination on our results, we first plotted the estimated distribution from contamination against the estimates of hake DNA concentration in field samples (Fig. \ref{fig:contam.facet}). The observed hake DNA distributions were all substantially larger than estimated contamination distribution. So while contamination may be playing a minor role in determining the observed hake DNA, it is not the dominant factor determining the spatial or depth specific distribution of DNA. We cannot identify which samples have a significant contribution of contamination but there is no suggestion that the contamination has a spatial or depth-specific effect.

We take these lessons from our experience with contamination.


```{r fig.contam.hist,warning=FALSE, echo=FALSE,fig.cap="\\label{fig:contam.hist} Histogram of point estimates for hake DNA copies per L from the two types of field control samples (N=49).  Blue line shows estimated lognormal distribution from the observed hake copies per L." ,out.width='80%'}
print(Output.summary.qpcr$control.hist)
```

```{r fig.contam.facet, echo=FALSE,fig.cap="\\label{fig:contam.facet} Empirical histograms of estimated hake DNA copies per L for individual samples at six water depths (purple) and estimated distribution of contamination derived from control samples. Vertical dashed line shows 20 copies per L, the putative detection threshold for our methodology. ",out.width='100%'}
print(Output.summary.qpcr$depth.hist.facet)
```


### Inhibition

As part of our qPCR protocol [@ramon-laca2021PLOS], we included an internal positive control (IPC) of the reaction to account for PCR inhibition. Any sample that had a difference of more that 0.5 PCR cycle from the IPC was considered to be inhibited.  Inhibited samples were initially purified using an inhibitor removal column and thereafter diluted 1:2, 1:5 and 1:10 to circumvent inhibition and requantified. We detected inhibition primarily in the near-surface samples (Fig. \ref{fig:inhibition.depth}).

```{r fig.inhibition.depth , echo=FALSE,fig.cap="\\label{fig:inhibition.depth} Samples that were identified as being inhibited by sample depths.  Samples identified as being inhibited were diluted and re-run using quantitative PCR. ",out.width='100%'}
print(Output.summary.qpcr$inhibit.plot.by.depth)
```

### Ethanol wash error

For a subset of samples, there was a laboratory error during the final ethanol wash following DNA extraction. Specifically, samples were washed with a 30\% ethanol solution instead of a 70% ethanol solution. Affected samples lost of some DNA during the washing processes. Washed bottles were geographically restricted to the northern part of the sample area, but occurred across all sample depths (Fig. \ref{fig:wash.plot}). To examine the magnitude of this problem, we experimentally divided samples taken from 26 individual additional station-depth combinations and subjected one bottle taken from those stations to the proper 70% ethanol wash and for the paired bottle used a 30% ethanol wash.  This paired design allowed us to estimate the magnitude of the DNA concentration lost due to the wash error within the statistical model.

```{r fig.wash.plot , echo=FALSE,fig.cap="\\label{fig:wash.plot} Samples that were identified as being inhibited by sample depths.  Samples identified as being inhibited were diluted and re-run using quantitative PCR. ",out.width='75%'}
print(Output.summary.qpcr$wash.plot)
```

We found that the ethanol was 

Output.summary.qpcr$wa


\newpage
\clearpage


### Scales of variation


```{r fig.bottle.var.depth , echo=FALSE,fig.cap="\\label{fig:bottle.var.depth} Estimates of TAU (log10 standard deviation of hake eDNA concentration between bottles at a single station-depth) for each sampling depth (posterior mean and 90 CI shown). Bottles at a sample location become increasingly similar at deeper sampling depths ",out.width='75%'}
print(Output.summary.qpcr$tau.sample.by.depth)
```

<!--Output.summary.qpcr$tau_long_summary -->

Output.summary.qpcr$tau.sample.by.depth


### CV maps for all methods

### D_Delta plots (roughly raw data)

### Marginal abundance with depth plots.

### Components of the acoustic model - pres-abs, positive, unconditional

### Raw predicted-observed plots for both models.

### Posterior variance for acoustics, eDNA.
