---
title: "Eulachon WCBTS index"
author: "Owen R. Liu"
date: "2023-11-21"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sdmTMB)
library(nwfscSurvey)
library(terra)
library(tidyverse)
library(lubridate)
library(rnaturalearth)
library(sf)
library(viridis)
library(here)
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)

plot_theme <- theme_minimal()+theme(panel.border = element_rect(color='black',fill=NA))
theme_set(plot_theme)
```

# Purpose

Set up a quick sdmTMB-based index of spatial abundance for eulachon, utilizing WCBTS data.

# Find Data

Download eulachon data from the FRAM data warehouse

```{r}
# pull the data
trawlraw <- PullCatch.fn(Name="eulachon",SurveyName = "NWFSC.Combo")
# organize and rename some variables
trawl <- trawlraw %>% 
  # convert date from character to date
  mutate(date=as_date(Date),
         Trawl_id=as.numeric(Trawl_id)) %>% 
  # clean up some other columns to be consistent
  mutate(species=tolower(Common_name)) %>% 
  rename(year=Year,depth_trawl=Depth_m) %>% 
  dplyr::select(year,date,depth_trawl,lon=Longitude_dd,lat=Latitude_dd,cpue_kg_km2)
```
## Distribution of Samples

```{r}
# crop to general areas that have had positive observations at some point in the past
trawl_crop <- trawl %>% 
  filter(cpue_kg_km2>0)
# crop to areas <= 500m depth?
trawl_depthcrop <- trawl %>% 
  filter(depth_trawl<=500)
trawl %>%
  sample_n(5000) %>% 
  ggplot(aes(lon,lat))+
  geom_point(size=0.5)+
  coord_equal()+
  ggtitle("Extent of All Samples")

trawl_crop %>%
  ggplot(aes(lon,lat))+
  geom_point(size=0.5)+
  coord_equal()+
  ggtitle("Positive Samples")

trawl_depthcrop%>%
  ggplot(aes(lon,lat))+
  geom_point(size=0.5)+
  coord_equal()+
  ggtitle("Samples <= 500m")
```


# Model Setup

Set up the parameters and options for sdmTMB

## 5km Grid

```{r}
# Blake's raster with just grid cell ID numbers
dat_raster=rast(here('Data','raster_grid_blake','fivekm_grid.tif'))

# Depth associated with grid cell IDs
raster_depth <- read_csv(here('Data','raster_grid_blake','weighted_mean_NGDC_depths_for_5km_gridcells.csv')) %>% 
  rename(depth_m= WM_depth_m) %>% 
  # make depth positive
  mutate(depth_m= -depth_m)

# join depths to cell numbers
cellnums <- tibble(rastID=as.numeric(values(dat_raster))) %>% 
  mutate(rastcell=row_number()) %>% 
  left_join(raster_depth,by=c("rastID"="Gridcell_ID"))

# make bathy raster
rast_depth <- setValues(dat_raster,cellnums$depth_m)
plot(rast_depth)
```
## Prediction Grid

Use the depth-cropped trawl data to make a prediction grid
```{r}
# convert trawl data to sf and project to Blake's CRS
trawlsf <- trawl_depthcrop %>% 
  st_as_sf(coords=c("lon","lat"),crs=4326) %>% 
  st_transform(crs(dat_raster))
# add projected coords to original data
trawlfit <- trawl_depthcrop %>% 
  mutate(x=st_coordinates(trawlsf)[,1],y=st_coordinates(trawlsf)[,2]) %>% 
  mutate(xkm=x/1000,ykm=y/1000)

# as a tibble for prediction
grid.pred <- as.data.frame(rast_depth,xy=T) %>% 
  mutate(depth_trawl=fivekm_grid)
# reduce to a bounding box that matches the mesh
predbbox <- trawlsf %>%
  summarise() %>% 
  st_convex_hull()

grid.pred <- grid.pred %>% 
  st_as_sf(coords=c("x","y"),crs=crs(dat_raster),remove = F) %>% 
  st_intersection(predbbox) %>% 
  mutate(xkm=x/1000,ykm=y/1000) %>% 
  st_set_geometry(NULL)

ggplot(grid.pred)+
  geom_point(aes(xkm,ykm),size=0.5)+
  coord_equal()
```

```{r}
# Make spde
spde <- make_mesh(trawlfit,xy_cols = c('xkm','ykm'),cutoff=5)
# model formula
formula <- "cpue_kg_km2 ~ s(depth_trawl)"
```
# Model: No Temporal Variation
Fit a model for the eulachon data with spatial RFs turned on. However, we don't deal with time/spatiotemporal variation (yet).

## Fit

```{r}
fit <- sdmTMB(
  formula=as.formula(formula),
  data = trawlfit,
  mesh = spde,
  family = tweedie(link = "log"),
  spatial = "on"
)
# some summaries
fit
tidy(fit,conf.int = T)
tidy(fit, effects = "ran_pars", conf.int = TRUE)
sanity(fit)
```


## Predict

```{r}
# coastline background map
coast <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>%
  st_transform(crs = crs(dat_raster))
```

```{r}
preds <- predict(fit,newdata=grid.pred)

glimpse(preds)

bbox <- st_bbox(predbbox)
preds.plot <- ggplot() + 
  geom_raster(data=preds, aes(x,y,fill = exp(est_non_rf)))+
  geom_sf(data=coast,col=NA,fill='gray80')+
  scale_fill_viridis()+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  # # scale_fill_viridis_b(breaks=c(0,0.01,0.1,0.2,0.3))+
  labs(x="",y="",fill="kg/km2",title="Eulachon density")
preds.plot
```

Marginal effect of depth

```{r}
ggeffects::ggpredict(fit, "depth_trawl [all]") |> plot()
```
# Model: Spatiotemporal Variation
Fit a model for the eulachon data with spatial RFs and S-T effects turned on, with an IID structure.

```{r}
fit_st <- sdmTMB(
  formula=as.formula(formula),
  data = trawlfit,
  mesh = spde,
  family = tweedie(link = "log"),
  spatial = "on",
  time='year',
  spatiotemporal = "iid"
)
# some summaries
fit_st
tidy(fit_st,conf.int = T)
sanity(fit_st)

write_rds(fit_st,here("Output","eulachon_trawl_sdm_st.rds"))
```

```{r}

```

