---
title: "Hake 2019 Survey"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE,warning=FALSE, include=FALSE}
source("post-process qPCR STAN Output.R")
```

## eDNA from the 2019 Hake cruise 

### Sampling and processing to date.

In total we have processed qPCR for pacific hake, pacific lamprey, and eulachon from `r N.station.run` CTD stations representing `r N.station.depth.run` station-depth combinations for `r N.sample.run` individual 2.5L water samples.   

```{r,echo=FALSE,warning=FALSE, fig.height=7}
print(sample.process.facet)
```


\newpage

### Inhibition

There are a lot of samples that show signs of inhibition. They are concentrated in the surface samples. Here are the inhibition by station-depths and depth categories.  'Completely inhibited' means zero PCR replicates from any water sample at that station-depth combination have amplified within the PCR parameters.  'Not inhibited' indicates that no PCR from that station-depth has evidence of inhibition. Insert language about non-template controls, how inhibition is identified. 

```{r,echo=FALSE}
kable(g_i_dat)
```

We can look at the patterns of inhibition spatially and by depth.

```{r,echo=FALSE,fig.height=8}
print(inhibit.plot.by.depth)
```

\clearpage
\newpage
Here are some larger plot of just a few shallow depths.  Inhibition is clearly strongest at the surface, but beyond depth there is no obvious pattern of inhibition among the samples.  Perhaps there is a hint of inhibition in the offsore edges of the transects at the 50 and 100m depths, but it is not a strong pattern.

```{r,echo=FALSE,fig.height=6}
print(inhibit.plot.by.depth.few)
```

\newpage
### Standards

The standards used in the qPCR assay look pretty reasonable good.  Each color represents a different PCR plate.

```{r,echo=FALSE,fig.height=3.5,fig.width=6,warning=FALSE}
print(stand.plot)
```

```{r,echo=FALSE,fig.height=3.5,fig.width=6,warning=FALSE}
print(stand.plot.pres)
```


\newpage
### Controls

There are two main types of control: field negative control and extraction controls.  The extraction controls have never produced any measurable amount of hake DNA, so we will ignore them for now.

Thus far we have analyzed `r N_control_run` samples out of a total of `r N_control_total` field negative samples. Out of the `r N_control_run`, `r N_control_positive` have detectable amounts of hake DNA. In general the amount of detected hake DNA is low  but is non-zero. 

```{r,echo=FALSE,fig.height=3,fig.width=5,warning=FALSE}
print(control.hist)
```

We can look at the temporal pattern of the control values too.  This will be more informative when we have more samples analyzed.

```{r,echo=FALSE,fig.height=3,fig.width=5,warning=FALSE}
print(control.by.time)
```

\newpage
## Hake eDNA

We can first ask if the distribution of estimated hake DNA compare to the values present in the field samples.  The water samples are substantially larger than the controls... which is good.

```{r,echo=FALSE,fig.height=3.5,fig.width=5.5,warning=FALSE}
print(compare.hist)
```

Let's make some plots of the spatial distribution of hake DNA.

```{r,echo=FALSE,fig.height=8,warning=FALSE}
print(hake.p2)
```


```{r,echo=FALSE,fig.height=9,warning=FALSE}
print(hake.p3)
```

```{r,echo=FALSE,fig.height=6,warning=FALSE}
print(depth.p2)
```

```{r,echo=FALSE,fig.height=6,warning=FALSE}
print(depth.p4)
```


