---
title: "Hake 2019 Survey"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## eDNA from the 2019 Hake cruise 

```{r pressure, echo=FALSE,warning=FALSE,include=FALSE}
library(tidyverse)
library(knitr)
library(reshape2)
library(viridis)

# Read in the output from the Stan model
base.dir <- "/Users/ole.shelton/Github/eDNA-Hake/Stan Model Fits"
setwd(base.dir)
load("qPCR Hake 2019 hake Fitted.RData")

### Get some of the output from file.
dat.samp <- Output.qpcr$dat.obs.bin
dat.control.field.neg <- Output.qpcr$dat.control.field.neg %>% left_join(.,Output.qpcr$field_neg_out)

dat.summary <- dat.samp %>%
                      group_by(date,transect,station,station.depth,depth,station_depth_idx,) %>%
                      dplyr::summarise(N=length(unique(sample_idx)))
dat.summary <- left_join(dat.summary,Output.qpcr$station_depth_out)
dat.summary <- dat.summary %>% mutate(depth_cat=case_when(depth < 25~0,
                                                          depth == 25~ 25,  
                                                          depth > 25 & depth <=50 ~ 50,
                                                          depth > 50 & depth <=100 ~ 100,
                                                          depth > 120 & depth <=150 ~ 150,
                                                          depth > 151 & depth <=200 ~ 200,
                                                          depth > 250 & depth <=350 ~ 300,
                                                          depth > 400 & depth <=500 ~ 500))

dat.inhibit <- Output.qpcr$dat.inhibit %>% mutate(depth_cat=case_when(depth < 25~ 0,
                                                          depth == 25~ 25,            
                                                          depth > 25 & depth <=50 ~ 50,
                                                          depth > 50 & depth <=100 ~ 100,
                                                          depth > 120 & depth <=150 ~ 150,
                                                          depth > 151 & depth <=200 ~ 200,
                                                          depth > 250 & depth <=350 ~ 300,
                                                          depth > 400 & depth <=500 ~ 500))

##### Calculate some info about the field negative controls
field.neg.summ <- Output.qpcr$field_neg_out 

N.station.run       <- length(unique(c(dat.inhibit$station,dat.samp$station)))
N.station.depth.run <- length(unique(c(dat.inhibit$station.depth,dat.samp$station.depth)))
N.sample.run        <- length(unique(c(dat.inhibit$sample,dat.samp$sample)))

```

### Sampling and processing to date.

In total we have processed qPCR for pacific hake, pacific lamprey, and eulachon from `r N.station.run` CTD stations representing `r N.station.depth.run` station-depth combinations for `r N.sample.run` individual 2.5L water samples.   

```{r, echo=FALSE,warning=FALSE,include=FALSE}
# Make a map of where we have samples.
library(ggmap)
library(maps)
library(mapdata)
library(ggplot2)

dat.id <- Output.qpcr$dat.id
dat.station.id.trim <- Output.qpcr$dat.station.id.trim
dat.sample.id <- Output.qpcr$dat.sample.id

# combine ids into stations, samples outstanding, and processed samples

dat.A <- dat.station.id.trim %>% dplyr::select(station,lat,lon) %>% mutate(Level="CTD Stations")
dat.B <- left_join(dat.sample.id,dat.id) %>% dplyr::select(station,lat,lon) %>%
            group_by(station,lat,lon) %>% summarise(N=length(station)) %>% filter(is.na(lat)==F) %>%
            mutate(Level="Sampled Stations")
dat.C <- dat.samp %>% dplyr::select(station,lat,lon) %>% group_by(station,lat,lon) %>% summarise(N=length(station)) %>%
            dplyr::select(-N) %>% mutate(Level="Processed Stations")


### THIS IS ALL OF THE CTD STATIONS.  IF YOU NEED A LAT LONG FOR EACH STATION, MERGE IN FROM HERE.
dat.process <- bind_rows(dat.A,dat.B,dat.C)
dat.process$Level <- factor(dat.process$Level, levels=c("CTD Stations","Sampled Stations","Processed Stations"))

dat.lat.lon <- dat.process %>% dplyr::select(station,lat,lon) %>% 
                  group_by(station) %>% summarise(mean.lat=mean(lat),mean.lon=mean(lon)) %>% 
                  filter(!station=="") %>% rename(lat=mean.lat,lon=mean.lon)
                          

#### Plotting information
states <- map_data("state")
west_coast <- subset(states, region %in% c("california", "oregon", "washington"))

# Plot for all samples
lat.lims <- c(min(dat.id$lat,na.rm=T),max(dat.id$lat,na.rm=T))
lon.lims <- c(min(dat.id$lon,na.rm=T),max(dat.id$lon,na.rm=T))

base_map <-ggplot(data = west_coast) + 
            geom_polygon(aes(x = long, y = lat, group=group), fill = grey(0.5), color = "black")+
            coord_fixed(xlim=lon.lims,ylim=lat.lims,ratio=1.3) +
            xlab("Longitude") +
            ylab("Latitude") +
            theme_bw()

sample.process <- base_map + 
                   scale_color_manual(values=c(grey(0.6),"blue","red"))+
                   geom_point(data=dat.process,aes(x=lon,y=lat,color=Level),size=0.8) 

sample.process.facet <- sample.process +
                        facet_wrap(~Level)

#print(sample.process)

```
<!---
```{r,echo=FALSE,warning=FALSE, fig.height=7}
print(sample.process)
```
--->

```{r,echo=FALSE,warning=FALSE, fig.height=7}
print(sample.process.facet)
```


\newpage

### Inhibition

```{r,echo=FALSE,warning=FALSE,results="asis",include=FALSE}
##### Calculate some inhibition characteristics.
inhibit.summ <- dat.inhibit %>% 
                      group_by(date,transect,station,station.depth,depth,depth_cat) %>%
                      summarise(N_pcr=length(unique(qPCR)),N_sample=length(unique(sample)),N_rep=length(sample)) %>%
                mutate(inhibited = "Y")
good.summ <- dat.summary %>%
                    dplyr::select(date,transect,station,station.depth,depth,depth_cat,N_sample=N) %>%
                    mutate(inhibited = "N")

g_i_dat <- bind_rows(inhibit.summ %>% dplyr::select(-N_pcr,-N_rep),good.summ) %>%
                  group_by(station,depth_cat,station.depth,inhibited) %>%
                  summarise(N_inhibited=length(inhibited)) %>%
                  pivot_wider(.,names_from=c("inhibited"),values_from=c("N_inhibited")) %>%
                  mutate(N=ifelse(is.na(N),0,N),
                         Y=ifelse(is.na(Y),0,Y),
                         Both=ifelse(N==1 & Y==1,1,0)) %>%
                  group_by(depth_cat) %>% 
                  summarise(Total=length(station.depth),
                            Not_inhibited=Total-sum(Y),
                            Partially_inhibited = sum(Both),
                            Completely_inhibited = Total-sum(N)) %>%
                  filter(is.na(depth_cat)==F)

g_i_dat <- bind_rows(g_i_dat %>% mutate(depth_cat=as.character(depth_cat)),
                     colSums(g_i_dat)%>%t()%>%as.data.frame()%>%mutate(depth_cat="TOTAL")) %>%
                      rename("Depth Category (m)"=depth_cat,)

g_i_2 <- bind_rows(inhibit.summ %>% dplyr::select(-N_pcr,-N_rep),good.summ) %>%
                  group_by(station,depth_cat,station.depth,inhibited) %>%
                  dplyr::summarise(N_inhibited=length(inhibited)) %>%
                  pivot_wider(.,names_from=c("inhibited"),values_from=c("N_inhibited")) %>%
                  mutate(N=ifelse(is.na(N),0,N),
                         Y=ifelse(is.na(Y),0,Y),
                         Both=ifelse(N==1 & Y==1,1,0)) %>%
                  left_join(.,dat.lat.lon)

  
  
  
#####
g_i_3 <- g_i_2 %>% mutate(inhib_cat=case_when(N==1 & Y==0 ~ "None",
                                             N==0 & Y==1 ~ "All",
                                             N==1 & Y==1 ~ "Partial"))
g_i_3$inhib_cat <- factor(g_i_3$inhib_cat,levels=c("All","Partial","None"))

### Make a spatial plot of inhibition

inhibit.plot <- base_map +
                    geom_point(data=g_i_3,aes(x=lon,y=lat,color=inhib_cat)) +
                    scale_color_viridis_d(begin=0,end=0.8)+
                    facet_wrap(~inhib_cat)
                  
inhibit.plot.by.depth <- base_map +
                    geom_point(data=g_i_3 %>% filter(is.na(depth_cat)==F),aes(x=lon,y=lat,color=inhib_cat),alpha=0.5) +
                    facet_wrap(~depth_cat,nrow = 2) +
                    scale_color_viridis_d("Inhibition\nCategory",begin=0,end=0.8) 


inhibit.plot.by.depth.few <- base_map +
                    geom_point(data=g_i_3 %>% filter(depth_cat %in% c(0,50,100)) ,aes(x=lon,y=lat,color=inhib_cat),alpha=0.5) +
                    facet_wrap(~depth_cat,nrow = 1) +
                    scale_color_viridis_d("Inhibition\nCategory",begin=0,end=0.8) 

```


There are a lot of samples that show signs of inhibition. They are concentrated in the surface samples. Here are the inhibition by station-depths and depth categories.  'Completely inhibited' means zero PCR replicates from any water sample at that station-depth combination have amplified within the PCR parameters.  'Not inhibited' indicates that no PCR from that station-depth has evidence of inhibition. Insert language about non-template controls, how inhibition is identified. 

```{r,echo=FALSE}
kable(g_i_dat)
```

We can look at the patterns of inhibition spatially and by depth.

```{r,echo=FALSE,fig.height=8}
print(inhibit.plot.by.depth)
```

\clearpage
\newpage
Here are some larger plot of just a few shallow depths.  Inhibition is clearly strongest at the surface, but beyond depth there is no obvious pattern of inhibition among the samples.  Perhaps there is a hint of inhibition in the offsore edges of the transects at the 50 and 100m depths, but it is not a strong pattern.

```{r,echo=FALSE,fig.height=6}
print(inhibit.plot.by.depth.few)
```

\newpage
### Standards

The standards used in the qPCR assay look pretty reasonable good.  Each color represents a different PCR plate.

```{r,echo=FALSE,warning=FALSE,results="asis",include=FALSE}

OFFSET <- Output.qpcr$OFFSET
pars <- Output.qpcr$samp
PCR <- Output.qpcr$PCR
dat.stand.pos <- Output.qpcr$dat.stand.pos
dat.stand.bin <- Output.qpcr$dat.stand.bin

B1 <- apply(pars$beta_0,2,mean)
B2 <- apply(pars$beta_1,2,mean)

P0 <- apply(pars$phi_0,2,mean)
P1 <- apply(pars$phi_1,2,mean)

V0 <-  mean(pars$sigma_stand_int)
#V1 <- mean(pars$sigma_stand_slope)
#V2 <- mean(pars$sigma_stand_slope2)

# Plot regression against Standard
X <- seq(0,5,length.out=1000) - OFFSET 
Y <- t(B1 + B2 %*% t(X ))

STAND.REG <- data.frame(X=X,Y=Y)
STAND.REG <- melt(STAND.REG,id.vars="X",value.name="Y")

x.lim=c(min(X),max(X))
y.lim=c(20,40)


lab.temp<- data.frame(variable=paste0("Y.",PCR$plate_idx), qPCR=PCR$qPCR)
STAND.REG <-left_join(STAND.REG,lab.temp)

stand.plot <- ggplot(dat.stand.pos) +
  geom_point(aes(x=log10_copies - OFFSET ,y=Ct,color=qPCR),alpha=0.5) +
  scale_shape_discrete(name="qPCR Plate") +
  theme_bw()
stand.plot <- stand.plot +
  geom_line(data=STAND.REG,aes(x=X,y=Y,color=qPCR),alpha=0.5) +
  scale_color_discrete(name="qPCR Plate") +
  ylab("PCR cycle")  +
  xlab("log10 copies DNA") 
#scale_x_continuous(labels = paste0("1e",LABS),limits = c(-2,4))
stand.plot

stand.plot.facet <- stand.plot +facet_wrap(~qPCR,ncol=2)

# Plot occurrence of standard
Y <- t(P0 + P1 %*% t(X ))
LOGIT <- data.frame(X=X,Y=plogis(Y))
LOGIT <- melt(LOGIT,id.vars="X",value.name="Y")

LOGIT <-left_join(LOGIT,lab.temp)

stand.plot.pres <- ggplot(dat.stand.bin) +
  geom_jitter(aes(x=log10_copies - OFFSET,y=Ct_bin,color=qPCR),alpha=0.75,width=0,height=0.05) +
  geom_line(data=LOGIT,aes(y=Y,x=X,color=qPCR)) +
  theme_bw() +
  scale_shape_discrete(name="qPCR Plate") +
  scale_color_discrete(name="qPCR plate") +
  xlab("log10 copies DNA") +
  ylab("Amplification success") +
  scale_y_continuous(breaks=c(0,1),labels = c("No","Yes"))
#scale_x_continuous(breaks=BREAKS,labels = paste0("1e",LABS),limits = c(-2.5,4))

stand.plot.pres

```

```{r,echo=FALSE,fig.height=3.5,fig.width=6,warning=FALSE}
print(stand.plot)
```

```{r,echo=FALSE,fig.height=3.5,fig.width=6,warning=FALSE}
print(stand.plot.pres)
```


\newpage
### Controls

```{r,echo=FALSE,warning=FALSE,results="asis",include=FALSE}
# Calculations for Field Negative Controls.
SAMPLES.CONTROL <- Output.qpcr$SAMPLES.CONTROL
dat.id.control <-  Output.qpcr$dat.id.control
field.neg.summ <- left_join(SAMPLES.CONTROL,field.neg.summ) %>% left_join(.,dat.id.control)

BREAKS <- seq(0,30,by=0.1)
XLIM <- c(0,max(field.neg.summ$Mean)*1.1)
control.hist <- ggplot(field.neg.summ) +
                  geom_histogram(aes(Mean),breaks=BREAKS) +
                  xlab("Estimated copies per uL") +
                  ylab("Number of negative controls") +
                  xlim(XLIM) +
                  theme_bw()
                  
ALPHA = 0.5 
control.by.time <- ggplot(field.neg.summ) +
                  geom_point(aes(x=date,y=Mean),alpha=ALPHA) +
                  geom_errorbar(aes(x=date,ymin=Val.X5.,ymax=Val.X95.),alpha=ALPHA,width=0) +
                  xlab("Date") +
                  ylab("Copies per uL") +
                  theme_bw()


N_control_positive <- field.neg.summ %>% filter(Mean>0.10) %>% nrow(.)
N_control_run <- nrow(field.neg.summ)
N_control_total <- nrow(dat.control.field.neg)

```

There are two main types of control: field negative control and extraction controls.  The extraction controls have never produced any measurable amount of hake DNA, so we will ignore them for now.

Thus far we have analyzed `r N_control_run` samples out of a total of `r N_control_total` field negative samples. Out of the `r N_control_run`, `r N_control_positive` have detectable amounts of hake DNA. In general the amount of detected hake DNA is low  but is non-zero. 

```{r,echo=FALSE,fig.height=3,fig.width=5,warning=FALSE}
print(control.hist)
```

We can look at the temporal pattern of the control values too.  This will be more informative when we have more samples analyzed.

```{r,echo=FALSE,fig.height=3,fig.width=5,warning=FALSE}
print(control.by.time)
```

\newpage
## Hake eDNA

We can first ask if the distribution of estimated hake DNA compare to the values present in the field samples.  The water samples are substantially larger than the controls... which is good.

```{r,echo=FALSE,warning=FALSE,results="asis",include=FALSE}

dat.sum.v.control <- bind_rows(dat.summary %>% dplyr::select(date,
                                                   Mean.log,Log.Val.X5.,Log.Val.X95.,
                                                   Mean,Val.X5.,Val.X95.) %>% mutate(cat="field") %>% as.data.frame(),
                               field.neg.summ %>% dplyr::select(date,
                                                   Mean.log,Log.Val.X5.,Log.Val.X95.,
                                                   Mean,Val.X5.,Val.X95.) %>% mutate(cat="control") %>% as.data.frame()
                                )



BREAKS <- 10^(c(-4,seq(-2,3,by=0.25)))
XLIM <- c(0,max(dat.sum.v.control$Mean)*1.01)
compare.hist <- ggplot(dat.sum.v.control) +
                  geom_histogram(aes(Mean,fill=cat),breaks=BREAKS,alpha=0.5,position="identity",) +
                  xlab("Estimated copies per uL") +
                  ylab("Number of Samples") +
                  xlim(XLIM) +
                  scale_fill_viridis_d("Category",begin=0,end=0.5,) +
                  scale_x_log10() +
                  theme_bw()
compare.hist

```

```{r,echo=FALSE,fig.height=3.5,fig.width=5.5,warning=FALSE}
print(compare.hist)
```


